### 2017-12-13 submission/manuscript.Rmd
change title
fix polymerase statement and add first paragraph discussion

# What the Taq? The Influence of Different Hi-Fidelity Taq Polymerase on 16S rRNA Sequencing

@@

***The Number of OTUs is Dependent on HiFi Taq Used:***
After normalization by individual, for each cycle number, we observed that for fecal samples the number of OTUs identified was dependent upon the HiFi Taq used and this difference increased as the depth of sub-sampling increased [Figure 1]. Lower cycle numbers (15-20) resulted in less differences between Taq while cycle numbers of 25, 30, and 35 had larger clearer defined differences [Figure 1]. Only 35 cycles had HiFi Taq differences that were significantly different at all sub-sampling levels (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. At sub-sampling depth of 5000 or higher 25 and 30 cycles had HiFi Taq differences (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05 & max(filter(fs_anova_summary, cycle == "25x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1]. Using a Tukey post-hoc test only at 35 cycles were significant differences found to be mainly driven by Platinum being different than all other Taq across sub-sampling depth (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2].

This Taq dependent difference in the number of OTUs was also observed in the mock community samples with the same Taq polymerases being high (Platinum) and low (Accuprime) respectively [Figure 2 & Table S3]. Conversely, differences between HiFi Taq were observed as early as 20 cycles and a sub-sampling depth of 1000 sequences (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. Using a Tukey post-hoc test differences between Platinum and the other HiFi Taqs was the major driver of the differences seen at different cycle numbers and sub-sampling depths [Table S4]. Both fecal and mock samples consistently showed that across sub-sampling depth and cycle number the lowest number of OTUs identified was from Accuprime^TM^ while the highest was from Platinum for both fecal and mock samples [Figure 1 & 2].

***Sequence Error is Dependent on both Taq and Cycle Number Used:***
Differences by HiFi Taq in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest values were for the Kappa HiFI Taq [Figure 3A]. Sub-sampling depth seems to have little effect on this rate with both 5000 and 10000 sub-sampled sequences showing similar results [Figure 3B-C]. There were small differences between the various HiFi Taq at lower cycle number but larger differences at higher cycle number with Platinum having large differences between all other HiFi Taq [Figure 3B-C and Table S6].

The total seqeunces with at least one error had multiple differences at different cycle numbers and sub-sampling depth driven by large differences in Accuprime^TM^ and Platinum versus the other HiFi Taq tested [Figure S1 & Table S7 & S8]. Although Accuprime^TM^ had the lowest per base error rate it had either the largest or second largest number of sequences with at least one error regardless of cycle number or sub-sampling depth [Figure S1]. Investigation of whether there were Taq dependent effects on base subsitution found that there was no clear bias and this was indpendent of sub-sampling depth [Figure S2-S4]. Further, the variation in subsitution error seems to reduce as the sub-sampling depth increases [Figure S2-S4].

***Chimeric Sequences Corelate with OTUs and are HiFi Taq Dependent:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi Taq used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Using a Dunn's post-hoc test the vast majority of these differences were driven by Platinum being different then all other HiFi Taq across cycle number and sub-sampling depth [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi Taq [Figure 4].

For all Taqs a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Platinum and Phusion HiFi Taq [Figure 5]. In general, the R^2^ value between the number of OTUs and chimeric sequences became stronger as sub-sampling depth increased [Figure 5]. Taken together this data suggests that a strong correlation exists between the number of OTUs and chimera sequence prevalence. Interestingly, Accuprime^TM^ not only had one of the lowest prevalence of chimeric sequences but also, consistently, had the lowest correlations between the number of OTUs and chimeric sequences across sub-sampling depth [Figure 5].

@@

>>>>

# What the Taq? The Influence of Different Hi-Fidelity Taq Polymerase on 16S rRNA Gene Sequencing

@@

***The Number of OTUs is Dependent on HiFi Taq Used:***
After normalization by individual, for each cycle number, we observed that for fecal samples the number of OTUs identified was dependent upon the HiFi DNA polymerase used and this difference increased as the depth of sub-sampling increased [Figure 1]. Lower cycle numbers (15-20) resulted in less differences between HiFI DNA polymerase while cycle numbers of 25, 30, and 35 had larger clearer defined differences [Figure 1]. Only 35 cycles had HiFi Taq differences that were significantly different at all sub-sampling levels (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. At sub-sampling depth of 5000 or higher 25 and 30 cycles had HiFi DNA polymerase differences (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05 & max(filter(fs_anova_summary, cycle == "25x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1]. Using a Tukey post-hoc test only at 35 cycles were significant differences found to be mainly driven by Platinum being different than all other Taq across sub-sampling depth (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2].

This HiFi DNA polymerase dependent difference in the number of OTUs was also observed in the mock community samples with the same DNA polymerases being high (Platinum) and low (Accuprime) respectively [Figure 2 & Table S3]. Conversely, differences between HiFi DNA polymerase were observed as early as 20 cycles and a sub-sampling depth of 1000 sequences (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. Using a Tukey post-hoc test differences between Platinum and the other HiFi DNA polymerases was the major driver of the differences seen at different cycle numbers and sub-sampling depths [Table S4]. Both fecal and mock samples consistently showed that across sub-sampling depth and cycle number the lowest number of OTUs identified was from Accuprime^TM^ while the highest was from Platinum for both fecal and mock samples [Figure 1 & 2].

***Sequence Error is Dependent on both Taq and Cycle Number Used:***
Differences by HiFi Taq in the median average per base error varied without a clear pattern across sub-sampling depth [Table S5]. Generally, the highest values were for the Kappa HiFI DNA polymerase [Figure 3A]. Sub-sampling depth seems to have little effect on this rate with both 5000 and 10000 sub-sampled sequences showing similar results [Figure 3B-C]. There were small differences between the various HiFi Taq at lower cycle number but larger differences at higher cycle number with Platinum having large differences between all other HiFi DNA polymerase [Figure 3B-C and Table S6].

The total seqeunces with at least one error had multiple differences at different cycle numbers and sub-sampling depth driven by large differences in Accuprime^TM^ and Platinum versus the other HiFi Taq tested [Figure S1 & Table S7 & S8]. Although Accuprime^TM^ had the lowest per base error rate it had either the largest or second largest number of sequences with at least one error regardless of cycle number or sub-sampling depth [Figure S1]. Investigation of whether there were HiFi DNA polymerase dependent effects on base subsitution found that there was no clear bias and this was indpendent of sub-sampling depth [Figure S2-S4]. Further, the variation in subsitution error seems to reduce as the sub-sampling depth increases [Figure S2-S4].

***Chimeric Sequences Corelate with OTUs and are HiFi Taq Dependent:***
After chimera removal using VSEARCH and removal of sequences that did not classify as bacteria we assessed the percentage of sequences that were still chimeric within our mock community. At all levels of sub-sampling and cycle number there were significant differences between the HiFi DNA polymerase used (P-value `r ifelse(max(filter(kruskal_chimera_prev, sub_sample_level > 500)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S9]. Using a Dunn's post-hoc test the vast majority of these differences were driven by Platinum being different then all other HiFi DNA polymerase across cycle number and sub-sampling depth [Table S10]. Generally, across sub-sampling depth and cycle number Accuprime^TM^ had the lowest chimera prevalence of all the HiFi DNA Polymerase [Figure 4].

For all Taqs a positive correlation was observed between chimeric sequences and number of OTUs, with this correlation being strongest for Platinum and Phusion HiFi DNA Polymerase [Figure 5]. In general, the R^2^ value between the number of OTUs and chimeric sequences became stronger as sub-sampling depth increased [Figure 5]. Taken together this data suggests that a strong correlation exists between the number of OTUs and chimera sequence prevalence. Interestingly, Accuprime^TM^ not only had one of the lowest prevalence of chimeric sequences but also, consistently, had the lowest correlations between the number of OTUs and chimeric sequences across sub-sampling depth [Figure 5].

@@

Our observations build upon previous studies [@taq_Gohl_2016] by showing that even different HiFi DNA Polymerase have significant differences in the number of OTUs and that the changes to total OTUs correlate with chimeras not removed after sequence processing [Figure 1-2 & 5]. This is important since many diversity metrics rely on the total number of OTUs as part of their calculations and changes to the total number of OTUs could drastically change the results as well as the findings. Although the attention has mostly been on standardizing and improving collection and extraction methods [@contamination_Salter2014] our observations show that independent of this consideration HiFi DNA polymerase can have a noticeable affect on the OTUs generated that can be found across sub-sampling depth and PCR cycle number [Figure 2-4].

```{r}
#Results have an impact on downstream diversity measurements due to effects on total OTUs

#Need to cover that different types of polymerase may affect outcome
#- Accuprime HiFi from recombinant A-Family Taq DNA polymerase, Pyrococcus spp GB-D polymerase (proof-reading), platinum Taq Antibody (inhibits room temp activity)

#- Kappa HiFi from B-family DNA polymerase (proof-reading baked in), propietary antibody (inhibit room temp activity)

#- Q5 HiFi from B-family thermostable DNA polymerase with 3´→ 5´ exonuclease activity, fused to a processivity-enhancing Sso7d domain to support robust DNA amplification

#-Phusion HiFi B-family DNA polymerase a novel Pyrococcus-like enzyme fused with a processivity-enhancing domain, generate blunt-ended products.

#- Platinum HiFi recombinant A-Family Taq DNA polymerase, Pyrococcus species GB-D polymeras, Platinum Taq Antibody

### 2017-12-13 submission/manuscript.tex
fix polymerase statement and add first paragraph discussion

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\textbf{\emph{The Number of OTUs is Dependent on HiFi Taq Used:}} After
normalization by individual, for each cycle number, we observed that for
fecal samples the number of OTUs identified was dependent upon the HiFi
Taq used and this difference increased as the depth of sub-sampling
increased {[}Figure 1{]}. Lower cycle numbers (15-20) resulted in less
differences between Taq while cycle numbers of 25, 30, and 35 had larger
clearer defined differences {[}Figure 1{]}. Only 35 cycles had HiFi Taq
differences that were significantly different at all sub-sampling levels
(P-value \textless{} 0.0001) {[}Table S1{]}. At sub-sampling depth of
5000 or higher 25 and 30 cycles had HiFi Taq differences (P-value
\textless{} 0.05) {[}Table S1{]}. Using a Tukey post-hoc test only at 35
cycles were significant differences found to be mainly driven by
Platinum being different than all other Taq across sub-sampling depth
(P-value \textless{} 0.05) {[}Table S2{]}.

This Taq dependent difference in the number of OTUs was also observed in
the mock community samples with the same Taq polymerases being high
(Platinum) and low (Accuprime) respectively {[}Figure 2 \& Table S3{]}.
Conversely, differences between HiFi Taq were observed as early as 20
cycles and a sub-sampling depth of 1000 sequences (P-value = 0.002)
{[}Table S3{]}. Using a Tukey post-hoc test differences between Platinum
and the other HiFi Taqs was the major driver of the differences seen at
different cycle numbers and sub-sampling depths {[}Table S4{]}. Both
fecal and mock samples consistently showed that across sub-sampling
depth and cycle number the lowest number of OTUs identified was from
Accuprime\textsuperscript{TM} while the highest was from Platinum for
both fecal and mock samples {[}Figure 1 \& 2{]}.

\textbf{\emph{Sequence Error is Dependent on both Taq and Cycle Number
Used:}} Differences by HiFi Taq in the median average per base error
varied without a clear pattern across sub-sampling depth {[}Table S5{]}.
Generally, the highest values were for the Kappa HiFI Taq {[}Figure
3A{]}. Sub-sampling depth seems to have little effect on this rate with
both 5000 and 10000 sub-sampled sequences showing similar results
{[}Figure 3B-C{]}. There were small differences between the various HiFi
Taq at lower cycle number but larger differences at higher cycle number
with Platinum having large differences between all other HiFi Taq
{[}Figure 3B-C and Table S6{]}.

The total seqeunces with at least one error had multiple differences at
different cycle numbers and sub-sampling depth driven by large
differences in Accuprime\textsuperscript{TM} and Platinum versus the
other HiFi Taq tested {[}Figure S1 \& Table S7 \& S8{]}. Although
Accuprime\textsuperscript{TM} had the lowest per base error rate it had
either the largest or second largest number of sequences with at least
one error regardless of cycle number or sub-sampling depth {[}Figure
S1{]}. Investigation of whether there were Taq dependent effects on base
subsitution found that there was no clear bias and this was indpendent
of sub-sampling depth {[}Figure S2-S4{]}. Further, the variation in
subsitution error seems to reduce as the sub-sampling depth increases
{[}Figure S2-S4{]}.

\textbf{\emph{Chimeric Sequences Corelate with OTUs and are HiFi Taq
Dependent:}} After chimera removal using VSEARCH and removal of
sequences that did not classify as bacteria we assessed the percentage
of sequences that were still chimeric within our mock community. At all
levels of sub-sampling and cycle number there were significant
differences between the HiFi Taq used (P-value \textless{} 0.05)
{[}Table S9{]}. Using a Dunn's post-hoc test the vast majority of these
differences were driven by Platinum being different then all other HiFi
Taq across cycle number and sub-sampling depth {[}Table S10{]}.
Generally, across sub-sampling depth and cycle number
Accuprime\textsuperscript{TM} had the lowest chimera prevalence of all
the HiFi Taq {[}Figure 4{]}.

For all Taqs a positive correlation was observed between chimeric
sequences and number of OTUs, with this correlation being strongest for
Platinum and Phusion HiFi Taq {[}Figure 5{]}. In general, the
R\textsuperscript{2} value between the number of OTUs and chimeric
sequences became stronger as sub-sampling depth increased {[}Figure
5{]}. Taken together this data suggests that a strong correlation exists
between the number of OTUs and chimera sequence prevalence.
Interestingly, Accuprime\textsuperscript{TM} not only had one of the
lowest prevalence of chimeric sequences but also, consistently, had the
lowest correlations between the number of OTUs and chimeric sequences
across sub-sampling depth {[}Figure 5{]}.

@@

@@

>>>>

\documentclass[12pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\textbf{\emph{The Number of OTUs is Dependent on HiFi Taq Used:}} After
normalization by individual, for each cycle number, we observed that for
fecal samples the number of OTUs identified was dependent upon the HiFi
DNA polymerase used and this difference increased as the depth of
sub-sampling increased {[}Figure 1{]}. Lower cycle numbers (15-20)
resulted in less differences between HiFI DNA polymerase while cycle
numbers of 25, 30, and 35 had larger clearer defined differences
{[}Figure 1{]}. Only 35 cycles had HiFi Taq differences that were
significantly different at all sub-sampling levels (P-value \textless{}
0.0001) {[}Table S1{]}. At sub-sampling depth of 5000 or higher 25 and
30 cycles had HiFi DNA polymerase differences (P-value \textless{} 0.05)
{[}Table S1{]}. Using a Tukey post-hoc test only at 35 cycles were
significant differences found to be mainly driven by Platinum being
different than all other Taq across sub-sampling depth (P-value
\textless{} 0.05) {[}Table S2{]}.

This HiFi DNA polymerase dependent difference in the number of OTUs was
also observed in the mock community samples with the same DNA
polymerases being high (Platinum) and low (Accuprime) respectively
{[}Figure 2 \& Table S3{]}. Conversely, differences between HiFi DNA
polymerase were observed as early as 20 cycles and a sub-sampling depth
of 1000 sequences (P-value = 0.002) {[}Table S3{]}. Using a Tukey
post-hoc test differences between Platinum and the other HiFi DNA
polymerases was the major driver of the differences seen at different
cycle numbers and sub-sampling depths {[}Table S4{]}. Both fecal and
mock samples consistently showed that across sub-sampling depth and
cycle number the lowest number of OTUs identified was from
Accuprime\textsuperscript{TM} while the highest was from Platinum for
both fecal and mock samples {[}Figure 1 \& 2{]}.

\textbf{\emph{Sequence Error is Dependent on both Taq and Cycle Number
Used:}} Differences by HiFi Taq in the median average per base error
varied without a clear pattern across sub-sampling depth {[}Table S5{]}.
Generally, the highest values were for the Kappa HiFI DNA polymerase
{[}Figure 3A{]}. Sub-sampling depth seems to have little effect on this
rate with both 5000 and 10000 sub-sampled sequences showing similar
results {[}Figure 3B-C{]}. There were small differences between the
various HiFi Taq at lower cycle number but larger differences at higher
cycle number with Platinum having large differences between all other
HiFi DNA polymerase {[}Figure 3B-C and Table S6{]}.

The total seqeunces with at least one error had multiple differences at
different cycle numbers and sub-sampling depth driven by large
differences in Accuprime\textsuperscript{TM} and Platinum versus the
other HiFi Taq tested {[}Figure S1 \& Table S7 \& S8{]}. Although
Accuprime\textsuperscript{TM} had the lowest per base error rate it had
either the largest or second largest number of sequences with at least
one error regardless of cycle number or sub-sampling depth {[}Figure
S1{]}. Investigation of whether there were HiFi DNA polymerase dependent
effects on base subsitution found that there was no clear bias and this
was indpendent of sub-sampling depth {[}Figure S2-S4{]}. Further, the
variation in subsitution error seems to reduce as the sub-sampling depth
increases {[}Figure S2-S4{]}.

\textbf{\emph{Chimeric Sequences Corelate with OTUs and are HiFi Taq
Dependent:}} After chimera removal using VSEARCH and removal of
sequences that did not classify as bacteria we assessed the percentage
of sequences that were still chimeric within our mock community. At all
levels of sub-sampling and cycle number there were significant
differences between the HiFi DNA polymerase used (P-value \textless{}
0.05) {[}Table S9{]}. Using a Dunn's post-hoc test the vast majority of
these differences were driven by Platinum being different then all other
HiFi DNA polymerase across cycle number and sub-sampling depth {[}Table
S10{]}. Generally, across sub-sampling depth and cycle number
Accuprime\textsuperscript{TM} had the lowest chimera prevalence of all
the HiFi DNA Polymerase {[}Figure 4{]}.

For all Taqs a positive correlation was observed between chimeric
sequences and number of OTUs, with this correlation being strongest for
Platinum and Phusion HiFi DNA Polymerase {[}Figure 5{]}. In general, the
R\textsuperscript{2} value between the number of OTUs and chimeric
sequences became stronger as sub-sampling depth increased {[}Figure
5{]}. Taken together this data suggests that a strong correlation exists
between the number of OTUs and chimera sequence prevalence.
Interestingly, Accuprime\textsuperscript{TM} not only had one of the
lowest prevalence of chimeric sequences but also, consistently, had the
lowest correlations between the number of OTUs and chimeric sequences
across sub-sampling depth {[}Figure 5{]}.

@@

Our observations build upon previous studies (Gohl et al., 2016) by
showing that even different HiFi DNA Polymerase have significant
differences in the number of OTUs and that the changes to total OTUs
correlate with chimeras not removed after sequence processing {[}Figure
1-2 \& 5{]}. This is important since many diversity metrics rely on the
total number of OTUs as part of their calculations and changes to the
total number of OTUs could drastically change the results as well as the
findings. Although the attention has mostly been on standardizing and
improving collection and extraction methods (Salter et al., 2014) our
observations show that independent of this consideration HiFi DNA
polymerase can have a noticeable affect on the OTUs generated that can
be found across sub-sampling depth and PCR cycle number {[}Figure
2-4{]}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{#Results have an impact on downstream diversity measurements due to effects on total OTUs}

\CommentTok{#Need to cover that different types of polymerase may affect outcome}
\CommentTok{#- Accuprime HiFi from recombinant A-Family Taq DNA polymerase, Pyrococcus spp GB-D polymerase (proof-reading), platinum Taq Antibody (inhibits room temp activity)}

\CommentTok{#- Kappa HiFi from B-family DNA polymerase (proof-reading baked in), propietary antibody (inhibit room temp activity)}

\CommentTok{#- Q5 HiFi from B-family thermostable DNA polymerase with 3´→ 5´ exonuclease activity, fused to a processivity-enhancing Sso7d domain to support robust DNA amplification}

\CommentTok{#-Phusion HiFi B-family DNA polymerase a novel Pyrococcus-like enzyme fused with a processivity-enhancing domain, generate blunt-ended products.}

\CommentTok{#- Platinum HiFi recombinant A-Family Taq DNA polymerase, Pyrococcus species GB-D polymeras, Platinum Taq Antibody}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{ref-contamination_Salter2014}{}
Salter SJ., Cox MJ., Turek EM., Calus ST., Cookson WO., Moffatt MF.,
Turner P., Parkhill J., Loman NJ., Walker AW. 2014. Reagent and
laboratory contamination can critically impact sequence-based microbiome
analyses. \emph{BMC Biology} 12. DOI:
\href{https://doi.org/10.1186/s12915-014-0087-z}{10.1186/s12915-014-0087-z}.
